### 2024-07-18 mathematics.tex
add discussion on Clifford Algebra

\detailstyle
\section{Mathematics}
\subsection{Matrix exponential}
Excerpted from \S2 and \S5 of Hall 2015 \cite{Hall2015}:
\begin{alignat}{2}
 &\ee^X := \sum_{m=0}^\infty \frac{X^m}{m!} \text{~~(converges for any $X$)},
\quad&
 &\log X := \sum_{m=1}^\infty (-1)^{m+1}\frac{(A-1)^m}{m} \text{~~(conv.~if $\|A-I\|<1$)}.
\end{alignat}
\begin{alignat}{2}
 &\ee^{\log A} = A \text{~~(if $\|A-I\|<1$)},
\quad&
 &\log \ee^X = X \text{~and~} \|\ee^X-1\| < 1 \text{~~(if $\|X\|<\log2$).}
\end{alignat}
\begin{equation}
 \text{Hilbert-Schmidt norm}:~\|X\|^2 := \sum_{i,j}|X_{ij}|^2 = {\Tr X^\dagger X}.
\end{equation}
Properties:
\begin{alignat*}{4}
 \ee^{(X^\TT)} &= (\ee^X)^\TT,
&\qquad
 \ee^{(X^*)} &= (\ee^X)^*,
&\qquad
 (\ee^X)^{-1} &= \ee^{-X},
&\qquad
 \ee^{YXY^{-1}} &= Y\ee^X Y^{-1},
\end{alignat*}
\begin{alignat*}{2}
 \det \exp X &= \exp\Tr X,
&\qquad
 \ee^{(\alpha+\beta)X} &= \ee^{\alpha X}\ee^{\beta X} \text{~for $\alpha, \beta\in \mathbb C$};
\end{alignat*}
Baker-Campbell-Hausdorff:
\begin{align}
  \ee^X Y \ee^{-X}
&= Y + [X, Y] + \frac1{2!}[X, [X, Y]] + \frac1{3!}[X, [X, [X, Y]]] + \cdots = \ee^{[X,]}Y;
\\
  \ee^X \ee^Y \ee^{-X}
&= \sum_{n=0}^\infty\frac{1}{n!}(\ee^X Y \ee^{-X})^n
 = \exp\left(\ee^{[X,]}Y\right);
\\
  \log(\ee^X \ee^Y)
&= X + \int_0^1\!\dd t\, g(\ee^{[X,}\ee^{t[Y,})Y
\qquad\Bigl[
  g(z) = \frac{\log z}{1-z^{-1}} = 1-\sum_{n=1}^{\infty}\frac{(1-z)^n}{n(n+1)};
\quad
g(\ee^y)=\sum_{n=0}^{\infty}\frac{B_ny^n}{n!}
\Bigr]\\
&= X + Y + \frac12[X,Y] + \frac1{12}[X, [X,Y]] - \frac1{12}[Y, [X,Y]] + \cdots
\quad\text{(Baker-Campbell-Hausdorff).}
\end{align}
\begin{align}
   \log(\ee^X\ee^Y)
&=\sum_{k=1}^\infty \frac{(-1)^{k-1}}{k}\left(\sum_{m,n=0}^{\infty}\frac{X^mY^n}{m!n!}-1\right)^k
&=\sum_{k=1}^\infty
   \sum_{m_1+n_1>0}\cdots\sum_{m_k+n_k>0}\frac{(-1)^{k-1}}{k}
\frac{X^{m_1}Y^{n_1}\cdots X^{m_k}Y^{n_k}}{m_1!n_1!\cdots m_k!n_k!}
\end{align}
\begin{equation}
   \log(\ee^X\ee^Y)=\sum_{k=1}^{\infty}\sum_{m_1+n_1>0}\cdots\sum_{m_k+n_k>0}
\frac{(-1)^{k-1}}{k\sum_{i=1}^{k}(m_i+n_i)}
\frac{
\Bigl([X,\Bigr)^{m_1}
\Bigl([Y,\Bigr)^{n_1}\cdots
\Bigl([X,\Bigr)^{m_k}
\Bigl([Y,\Bigr)^{n_k}]\cdots]
}{m_1!n_1!\cdots m_k!n_k!}
\end{equation}
\hfill with $[X]$ being $X$.
\\
Derivative:
\begin{alignat}{1}
 &\frac{\dd}{\dd t}\ee^{tX} = X\ee^{t X} = \ee^{tX} X\\
 &\ee^{-X(t)}\left(\frac{\dd}{\dd t}\ee^{X(t)}\right) =
    \frac{I-\ee^{-\ad_X}}{\ad_X}\left(\frac{\dd X}{\dd t}\right)
  = X' + \frac{[-X,X']}{2!} + \frac{[-X,[-X,X']]}{3!} + \cdots
\\
 &\left(\frac{\dd}{\dd t}\ee^{X(t)}\right)\ee^{-X(t)}
= X' + \frac{[X,X']}{2!} + \frac{[X,[X,X']]}{3!} + \cdots
\end{alignat}
\hfill where $X'=\dd X/\dd t$ and $\ad_X(Y)=[X,Y]$ is the adjoint action of a Lie algebra. Thus, explicitly,
\begin{alignat}{1}
\frac{\dd}{\dd t}\ee^{aX(t)}
  = \ee^{aX}\left\{\sum_{n=0}^\infty\frac{a^{n+1}}{(n+1)!}\Bigl([-X,\Bigr)^n X']\right\}
  = \left\{\sum_{n=0}^\infty\frac{a^{n+1}}{(n+1)!}\Bigl([X,\Bigr)^n X']\right\}\ee^{aX}
\end{alignat}

@@

\paragraph{Diagonalization}
A matrix $M\in\mathbb K^{n\times n}$ is called diagonalizable if $\exists P$ and $\exists D$ s.t.
\begin{equation}
 M=PDP^{-1};\qquad
 P\in\mathbb{K}^{n\times n},\quad
 D:\text{diagonal matrix}~(D_{ii}\in\mathbb{K}).
\end{equation}
In particular,
\begin{equation}
 \text{$M$ is normal} \stackrel{\text{def}}\iff M^\dagger M = M M^\dagger \iff
 \exists P\in\mathbb{U}_{\mathbb K}^{n} \text{~s.t.~} M=PDP^{-1}.
\end{equation}

\paragraph{Singular value decomposition}
Any $M\in\mathbb{K}^{m\times n}$ can be singular-value decomposed as
\begin{equation}
 M=UDV^\dagger;\qquad
 U\in\mathbb{U}_{\mathbb K}^m,\quad
 V\in\mathbb{U}_{\mathbb K}^n,\quad
 D: \text{non-negative real diagonal matrix}~(D_{ii}\ge0).
\end{equation}
Here, the matrix $U$ ($V$) diagonalizes $MM^\dagger$ ($M^\dagger M$) and $(D_{ii})^2$ are the eigenvalues of $MM^\dagger$ (and $M^\dagger M$).

@@

\paragraph{Autonne-Takagi factorization}
If $M\in\mathbb{C}^{n\times n}$ is symmetric, it can be decomposed as
\begin{equation}
 M=RDR^\TT;\qquad
 R\in\mathbb{U}_{\mathbb C}^n,\quad
 D: \text{non-negative real diagonal matrix}~(D_{ii}\ge0).
 \label{eq:ATF}
\end{equation}
Real symmetric matrices are normal and thus do not need this factorization; we can apply the above ``diagonalization'' method.

@@

With $\Omega=\spmat{0&I_{n/2}\\-I_{n/2}}$ and $H=\spmat{I_p&0\\0&-I_q}$,
\begin{align}
 \mathop{\mathrm  {GL}}(n;\KK) &= \{g\in M^n_\KK\mathbin|\det g\neq 0\},&
 \mathop{\mathfrak{gl}}(n;\KK) &= \{a\in M^n_\KK\},\\
%
 \mathop{\mathrm  {SL}}(n;\KK) &= \{g\in M^n_\KK\mathbin|\det g=1\},&
 \mathop{\mathfrak{sl}}(n;\KK) &= \{a\in M^n_\KK\mathbin|\Tr a=0\},\\
%
 \mathop{\mathrm  {SU}}(n) &= \{g\in M^n_{\mathbb C}\mathbin|g^\dagger g=1\land\det g=1\},&
 \mathop{\mathfrak{su}}(n) &= \{a\in M^n_{\mathbb C}\mathbin|a+a^\dagger=0\land\Tr a=0\},\\
%
 \mathop{\mathrm  {SO}}(n) &= \{g\in M^n_{\mathbb R}\mathbin|g^\TT g=1\land\det g=1\},&
 \mathop{\mathfrak{so}}(n) &= \{a\in M^n_{\mathbb R}\mathbin|a+a^\TT=0\},\\
%
 \mathop{\mathrm  {SO}}(n;\mathbb C) &= \{g\in M^n_{\mathbb C}\mathbin|g^\TT g=1\land\det g=1\},&
 \mathop{\mathfrak{so}}(n;\mathbb C) &= \{a\in M^n_{\mathbb C}\mathbin|a+a^\TT=0\},\\
%
 \mathop{\mathrm  {SO}}(p,q) &= \{g\in M^{p+q}_{\mathbb R}\mathbin|H g^\TT H=g^{-1}\land\det g=1\},&
 \mathop{\mathfrak{so}}(p,q) &= \{a\in M^{p+q}_{\mathbb R}\mathbin|H a^\TT H=-a\},\\
%
 \mathop{\mathrm  {Sp}}(n;\KK) &= \{g\in M^n_\KK\mathbin|\Omega g^\TT \Omega=-g^{-1}\},&
 \mathop{\mathfrak{sp}}(n;\KK) &= \{g\in M^n_\KK\mathbin|\Omega a^\TT \Omega=a\},\\
%
 \mathop{\mathrm  {Sp}}(n) &= \{g\in M^n_{\mathbb C}\mathbin|\epsilon g^\TT \epsilon=-g^{-1}\land g^\dagger g=1\},&
 \mathop{\mathfrak{sp}}(n) &= \{g\in M^n_{\mathbb C}\mathbin|\epsilon a^\TT \epsilon=a\land a^\dagger+a=0\}.
\end{align}
Of course, $\mathop{\mathrm{U}}(n)$ and $\mathop{\mathrm{O}}(n;\KK)$ (and corresponding Lie algebra) are obtained by removing the condition $\det g=1$ (and $\Tr a=0$).

@@

The complexification of a finite-dimensional \emph{real} vector space $V$ is defined by
\begin{equation}
 V_{\mathbb C}=V\otimes\mathbb C=\{v_1+\ii v_2\mathbin|v_1,v_2\in V\}\cong V+\ii V.
\end{equation}
Then, complexification of a Lie algebra $\mathfrak a$ of a matrix Lie group is defined as follows.
Since $\mathfrak a$ is a \emph{real} Lie algebra, it can be complexified to $\mathfrak a_{\mathbb C}$, where its bracket operation has a unique extension so that $\mathfrak a_{\mathbb C}$ is a \emph{complex} lie algebra.
The complex Lie algebra $\mathfrak a_{\mathbb C}$ is called the complexification of $\mathfrak a$.
In particular,
\begin{align}\notag
 &
 \mathop{\mathfrak{gl}}(n;\mathbb R)_{\mathbb C}
 \cong \mathop{\mathfrak{u}}(n)_{\mathbb C}
 \cong \mathop{\mathfrak{gl}}(n;\mathbb C),&
 &
 \mathop{\mathfrak{so}}(n)_{\mathbb C}
 \cong \mathop{\mathfrak{so}}(n;\mathbb C),\\
 &
 \mathop{\mathfrak{sl}}(n;\mathbb R)_{\mathbb C}
 \cong \mathop{\mathfrak{su}}(n)_{\mathbb C}
 \cong \mathop{\mathfrak{sl}}(n;\mathbb C),&
 &
 \mathop{\mathfrak{sp}}(n;\mathbb R)_{\mathbb C}
 \cong \mathop{\mathfrak{sp}}(n)_{\mathbb C}
 \cong \mathop{\mathfrak{sp}}(n;\mathbb C).
\end{align}
On the other hand, a complex simple Lie algebra $\mathfrak a$ can be, on the other hand, viewed as a real Lie algebra, whose (real) dimension is twice as the original (complex) dimension.
We describe such real Lie algebra by $\mathfrak a_{\mathbb R}$.

@@

\subsection{Mathematical Foundations for Spinors}
\paragraph{Quadratic form}
Let~\cite[\S6.3]{Jacobson} $V$ be an $n$-dimensional vector space over a field $K$. A quadratic form $Q$ on $V$ is defined by
\begin{equation}
\begin{split}
   Q\colon V\to K,\qquad
   &\forall k\in K, v\in V, Q(k v)=k^2Q(v),\\
   &B\colon (v,w)\mapsto Q(v+w)-Q(v)-Q(w)\text{~is linear in $v$ and $w$.}
\end{split}
\end{equation}
The bilinear form $B$ satisfies $B(x,x)=2Q(x)$. As long as the characteristic of $F$ is not two, $Q$ is determined by $B$.

Equivalently, $Q$ is characterized by a symmetric ``matrix'' $\hat Q$ such that $Q(v)=v^\TT \hat Q v$.
Focusing on $K=\RR$, $\hat Q$ is a real symmetric matrix and thus diagonalizable. Sylvester's law of inertia states that there exists a basis such that $\hat Q=\diag(1,1,\cdots,1,-1,-1,\cdots,-1)$, where the number of $+1$ and $-1$ are respectively $p$ and $q$.
The pair $(p,q)$ is unique and is called the signature of $Q$, where $\rank \hat Q=p+q$.

Quadratic form induces orthogonality: a basis $\{e_i\}$ of $V$ is called orthogonal [orthonormal] under $Q$ if
\begin{equation}
  B(e_i,e_j)\propto \delta_{ij}\qquad\Bigl[B(e_i,e_j)=c_i\delta_{ij}\text{~with~}c_i\in\{2,0,-2\}\Bigr]
\end{equation}
and $(V,Q)$ has an orthogonal basis as long as the characteristic of $K$ is not two.
If $K$ is a ``spin field'' such as $\RR$ or $\CC$ (or, precisely, $\forall \alpha\in K,\exists \beta\in K$ s.t. $\alpha=\beta^2$ or $-\beta^2$), $(V,Q)$ has an orthonormal basis.

\paragraph{Geometric algebra}
Let~\cite{0907.5356} $V$ be an $n$-dimensional vector space over a field $K$. Let $Q$ be a quadratic form on $V$.
Tensor algebra on $V$ is defined by
\begin{equation}
  T(V) = \bigoplus_{k} V^{\otimes k}
       = K\oplus V\oplus (V\otimes V)\oplus\cdots.
\end{equation}
It has an ideal $K_Q$ generated by $\{v\otimes v - Q(v)\mid v\in V\}$:
\begin{equation}
  K_Q = \left\{\sum_{i} a_i\otimes(v_i\otimes v_i-Q(v_i))\otimes b_i\relmiddle| v_i\in V, a_i,b_i \in T(V)\right\}.
\end{equation}
Then the Geometric algebra $\GeoA(V,Q)$ is defined by $T(V)/K_Q$. This is isomorphic to $\Cl(E, K, q_E)$ defined below.

The product in $\GeoA(V,Q)$ is written by $vw \deq [v\otimes w]=v\otimes w+K_q$. Particularly, $v^2 = Q(v)$ and $vw+wv=B(v,w)$.

\paragraph{Clifford Algebra}
For a finite set $X$, a commutative ring $R$ with unit,
the Clifford algebra $\Cl(X,R,r)$ is defined as the free $R$-module generated by the set $2^X$ of all subset of $X$.
Here, $r\colon X\to R$ is an arbitrary function thought of as a signature.

The Clifford algebra $\mathrm{Cl}$ has
addition $\mathrm{Cl}\times\mathrm{Cl}\to\mathrm{Cl}$,
scalar multiplication $R\times\mathrm{Cl}\to\mathrm{Cl}$, and
product
\begin{equation}
  \mathrm{Cl}\times\mathrm{Cl}\to\mathrm{Cl}\ni AB
  \deq \tau(A,B)A\setdiff B,
\end{equation}
where $A\setdiff B=(A\cup B)\setminus(B\cap A)$,
defined based on the following map $\tau\colon2^X\times2^X\to R$:
\begin{equation}
  \begin{split}
  &\tau(\{x\},\{x\}) = r(x)\quad\forall x\in X,&
  &\tau(A,B)\tau(A\setdiff B,C)=\tau(A,B\setdiff C)\tau(B,C)\quad\forall A,B,C\in 2^X,\\
  &\tau(\{y\},\{x\}) = -\tau(\{x\},\{y\})\quad\forall x,y\in X,x\neq y,&
  &\tau(\emptyset,A)=\tau(A,\emptyset)=1\quad\forall A\in2^X,\\
  &\tau(A,B)\in\{-1,1\}\quad \text{if}~A\cap B=\emptyset.
  \end{split}
\end{equation}

Several standard operations are defined: for $A\in2^X$ and $k=\binom{|A|}{2}$,
\begin{equation}
\text{Grade involution}\quad A^\star\deq(-1)^{|A|}A,\qquad
\text{Reversion}\quad A^\dagger\deq(-1)^{k}A,\qquad
\text{Clifford conjugate}\quad A^\mdlgwhtsquare\deq A^{\star\dagger}.
\end{equation}
The sign of grade involution (reversion) is negative iff $|A|$ mod 4 is 1 or 3 (2 or 3).

Let $(V,Q)$ be an $n$-dimensional vector space $V$ over a field $K$ equipped with a quadratic form $Q$. Let $E=\{e_i\}$ be an orthogonal basis of $(V,Q)$.
Then
\begin{equation}
  \Cl(E, K, Q|_E) \cong \GeoA(V,Q).
\end{equation}
The Clifford algebra $\Cl(X,R,r)$ has subspaces of $k$-vectors $\mathrm{Cl}^k$ and of even and odd vectors $\mathrm{Cl}^\pm$:
\[
\Cl(X,R,r)=\mathrm{Cl}^+\oplus\mathrm{Cl}^-=
\mathrm{Cl}^0\oplus\mathrm{Cl}^1\oplus\mathrm{Cl}^2\oplus\cdots\oplus\mathrm{Cl}^{|X|}.
\]
In particular, $\mathrm{Cl}^+$ is a subalgebra.

>>>>

\detailstyle
\section{Mathematics}
\subsection{Matrix exponential}
Excerpted from \S2 and \S5 of Hall 2015 \cite{Hall2015}:
\begin{alignat}{2}
 &\ee^X \deq \sum_{m=0}^\infty \frac{X^m}{m!} \text{~~(converges for any $X$)},
\quad&
 &\log X \deq \sum_{m=1}^\infty (-1)^{m+1}\frac{(A-1)^m}{m} \text{~~(conv.~if $\|A-I\|<1$)}.
\end{alignat}
\begin{alignat}{2}
 &\ee^{\log A} = A \text{~~(if $\|A-I\|<1$)},
\quad&
 &\log \ee^X = X \text{~and~} \|\ee^X-1\| < 1 \text{~~(if $\|X\|<\log2$).}
\end{alignat}
\begin{equation}
 \text{Hilbert-Schmidt norm:}~\|X\|^2 \deq \sum_{i,j}|X_{ij}|^2 = {\Tr X^\dagger X}.
\end{equation}
Properties:
\begin{alignat*}{4}
 \ee^{(X^\TT)} &= (\ee^X)^\TT,
&\qquad
 \ee^{(X^*)} &= (\ee^X)^*,
&\qquad
 (\ee^X)^{-1} &= \ee^{-X},
&\qquad
 \ee^{YXY^{-1}} &= Y\ee^X Y^{-1},
\end{alignat*}
\begin{alignat*}{2}
 \det \exp X &= \exp\Tr X,
&\qquad
 \ee^{(\alpha+\beta)X} &= \ee^{\alpha X}\ee^{\beta X} \text{~for $\alpha, \beta\in \mathbb C$};
\end{alignat*}
Baker-Campbell-Hausdorff:
\begin{align}
  \ee^X Y \ee^{-X}
&= Y + [X, Y] + \frac1{2!}[X, [X, Y]] + \frac1{3!}[X, [X, [X, Y]]] + \cdots = \ee^{[X,]}Y;
\\
  \ee^X \ee^Y \ee^{-X}
&= \sum_{n=0}^\infty\frac{1}{n!}(\ee^X Y \ee^{-X})^n
 = \exp\left(\ee^{[X,]}Y\right);
\\
  \log(\ee^X \ee^Y)
&= X + \int_0^1\!\dd t\, g(\ee^{[X,}\ee^{t[Y,})Y
\qquad\Bigl[
  g(z) = \frac{\log z}{1-z^{-1}} = 1-\sum_{n=1}^{\infty}\frac{(1-z)^n}{n(n+1)};
\quad
g(\ee^y)=\sum_{n=0}^{\infty}\frac{B_ny^n}{n!}
\Bigr]\\
&= X + Y + \frac12[X,Y] + \frac1{12}[X, [X,Y]] - \frac1{12}[Y, [X,Y]] + \cdots
\quad\text{(Baker-Campbell-Hausdorff).}
\end{align}
\begin{align}
   \log(\ee^X\ee^Y)
&=\sum_{k=1}^\infty \frac{(-1)^{k-1}}{k}\left(\sum_{m,n=0}^{\infty}\frac{X^mY^n}{m!n!}-1\right)^k
&=\sum_{k=1}^\infty
   \sum_{m_1+n_1>0}\cdots\sum_{m_k+n_k>0}\frac{(-1)^{k-1}}{k}
\frac{X^{m_1}Y^{n_1}\cdots X^{m_k}Y^{n_k}}{m_1!n_1!\cdots m_k!n_k!}
\end{align}
\begin{equation}
   \log(\ee^X\ee^Y)=\sum_{k=1}^{\infty}\sum_{m_1+n_1>0}\cdots\sum_{m_k+n_k>0}
\frac{(-1)^{k-1}}{k\sum_{i=1}^{k}(m_i+n_i)}
\frac{
\Bigl([X,\Bigr)^{m_1}
\Bigl([Y,\Bigr)^{n_1}\cdots
\Bigl([X,\Bigr)^{m_k}
\Bigl([Y,\Bigr)^{n_k}]\cdots]
}{m_1!n_1!\cdots m_k!n_k!}
\end{equation}
\hfill with $[X]$ being $X$.
\\
Derivative:
\begin{alignat}{1}
 &\frac{\dd}{\dd t}\ee^{tX} = X\ee^{t X} = \ee^{tX} X\\
 &\ee^{-X(t)}\left(\frac{\dd}{\dd t}\ee^{X(t)}\right) =
    \frac{I-\ee^{-\ad_X}}{\ad_X}\left(\frac{\dd X}{\dd t}\right)
  = X' + \frac{[-X,X']}{2!} + \frac{[-X,[-X,X']]}{3!} + \cdots
\\
 &\left(\frac{\dd}{\dd t}\ee^{X(t)}\right)\ee^{-X(t)}
= X' + \frac{[X,X']}{2!} + \frac{[X,[X,X']]}{3!} + \cdots
\end{alignat}
\hfill where $X'=\dd X/\dd t$ and $\ad_X(Y)=[X,Y]$ is the adjoint action of a Lie algebra. Thus, explicitly,
\begin{alignat}{1}
\frac{\dd}{\dd t}\ee^{aX(t)}
  = \ee^{aX}\left\{\sum_{n=0}^\infty\frac{a^{n+1}}{(n+1)!}\Bigl([-X,\Bigr)^n X']\right\}
  = \left\{\sum_{n=0}^\infty\frac{a^{n+1}}{(n+1)!}\Bigl([X,\Bigr)^n X']\right\}\ee^{aX}
\end{alignat}

@@

\paragraph{Diagonalization}
A matrix $M\in\mathbb K^{n\times n}$ is called diagonalizable if $\exists P$ and $\exists D$ s.t.
\begin{equation}
 M=PDP^{-1};\qquad
 P\in\mathbb{K}^{n\times n},\quad
 D\colon\text{diagonal matrix}~(D_{ii}\in\mathbb{K}).
\end{equation}
In particular,
\begin{equation}
 \text{$M$ is normal} \stackrel{\text{def}}\iff M^\dagger M = M M^\dagger \iff
 \exists P\in\mathbb{U}_{\mathbb K}^{n} \text{~s.t.~} M=PDP^{-1}.
\end{equation}

\paragraph{Singular value decomposition}
Any $M\in\mathbb{K}^{m\times n}$ can be singular-value decomposed as
\begin{equation}
 M=UDV^\dagger;\qquad
 U\in\mathbb{U}_{\mathbb K}^m,\quad
 V\in\mathbb{U}_{\mathbb K}^n,\quad
 D\colon \text{non-negative real diagonal matrix}~(D_{ii}\ge0).
\end{equation}
Here, the matrix $U$ ($V$) diagonalizes $MM^\dagger$ ($M^\dagger M$) and $(D_{ii})^2$ are the eigenvalues of $MM^\dagger$ (and $M^\dagger M$).

@@

\paragraph{Autonne-Takagi factorization}
If $M\in\mathbb{C}^{n\times n}$ is symmetric, it can be decomposed as
\begin{equation}
 M=RDR^\TT;\qquad
 R\in\mathbb{U}_{\mathbb C}^n,\quad
 D\colon \text{non-negative real diagonal matrix}~(D_{ii}\ge0).
 \label{eq:ATF}
\end{equation}
Real symmetric matrices are normal and thus do not need this factorization; we can apply the above ``diagonalization'' method.

@@

With $\Omega=\spmat{0&I_{n/2}\\-I_{n/2}}$ and $H=\spmat{I_p&0\\0&-I_q}$,
\begin{align}
 \mathop{\mathrm  {GL}}(n;\KK) &= \{g\in M^n_\KK\mathbin\mid \det g\neq 0\},&
 \mathop{\mathfrak{gl}}(n;\KK) &= \{a\in M^n_\KK\},\\
%
 \mathop{\mathrm  {SL}}(n;\KK) &= \{g\in M^n_\KK\mathbin\mid \det g=1\},&
 \mathop{\mathfrak{sl}}(n;\KK) &= \{a\in M^n_\KK\mathbin\mid \Tr a=0\},\\
%
 \mathop{\mathrm  {SU}}(n) &= \{g\in M^n_{\mathbb C}\mathbin\mid g^\dagger g=1\land\det g=1\},&
 \mathop{\mathfrak{su}}(n) &= \{a\in M^n_{\mathbb C}\mathbin\mid a+a^\dagger=0\land\Tr a=0\},\\
%
 \mathop{\mathrm  {SO}}(n) &= \{g\in M^n_{\mathbb R}\mathbin\mid g^\TT g=1\land\det g=1\},&
 \mathop{\mathfrak{so}}(n) &= \{a\in M^n_{\mathbb R}\mathbin\mid a+a^\TT=0\},\\
%
 \mathop{\mathrm  {SO}}(n;\mathbb C) &= \{g\in M^n_{\mathbb C}\mathbin\mid g^\TT g=1\land\det g=1\},&
 \mathop{\mathfrak{so}}(n;\mathbb C) &= \{a\in M^n_{\mathbb C}\mathbin\mid a+a^\TT=0\},\\
%
 \mathop{\mathrm  {SO}}(p,q) &= \{g\in M^{p+q}_{\mathbb R}\mathbin\mid H g^\TT H=g^{-1}\land\det g=1\},&
 \mathop{\mathfrak{so}}(p,q) &= \{a\in M^{p+q}_{\mathbb R}\mathbin\mid H a^\TT H=-a\},\\
%
 \mathop{\mathrm  {Sp}}(n;\KK) &= \{g\in M^n_\KK\mathbin\mid \Omega g^\TT \Omega=-g^{-1}\},&
 \mathop{\mathfrak{sp}}(n;\KK) &= \{g\in M^n_\KK\mathbin\mid \Omega a^\TT \Omega=a\},\\
%
 \mathop{\mathrm  {Sp}}(n) &= \{g\in M^n_{\mathbb C}\mathbin\mid \epsilon g^\TT \epsilon=-g^{-1}\land g^\dagger g=1\},&
 \mathop{\mathfrak{sp}}(n) &= \{g\in M^n_{\mathbb C}\mathbin\mid \epsilon a^\TT \epsilon=a\land a^\dagger+a=0\}.
\end{align}
Of course, $\mathop{\mathrm{U}}(n)$ and $\mathop{\mathrm{O}}(n;\KK)$ (and corresponding Lie algebra) are obtained by removing the condition $\det g=1$ (and $\Tr a=0$).

@@

The complexification of a finite-dimensional \emph{real} vector space $V$ is defined by
\begin{equation}
 V_{\mathbb C}=V\otimes\mathbb C=\{v_1+\ii v_2\mathbin\mid v_1,v_2\in V\}\cong V+\ii V.
\end{equation}
Then, complexification of a Lie algebra $\mathfrak a$ of a matrix Lie group is defined as follows.
Since $\mathfrak a$ is a \emph{real} Lie algebra, it can be complexified to $\mathfrak a_{\mathbb C}$, where its bracket operation has a unique extension so that $\mathfrak a_{\mathbb C}$ is a \emph{complex} lie algebra.
The complex Lie algebra $\mathfrak a_{\mathbb C}$ is called the complexification of $\mathfrak a$.
In particular,
\begin{align}\notag
 &
 \mathop{\mathfrak{gl}}(n;\mathbb R)_{\mathbb C}
 \cong \mathop{\mathfrak{u}}(n)_{\mathbb C}
 \cong \mathop{\mathfrak{gl}}(n;\mathbb C),&
 &
 \mathop{\mathfrak{so}}(n)_{\mathbb C}
 \cong \mathop{\mathfrak{so}}(n;\mathbb C),\\
 &
 \mathop{\mathfrak{sl}}(n;\mathbb R)_{\mathbb C}
 \cong \mathop{\mathfrak{su}}(n)_{\mathbb C}
 \cong \mathop{\mathfrak{sl}}(n;\mathbb C),&
 &
 \mathop{\mathfrak{sp}}(n;\mathbb R)_{\mathbb C}
 \cong \mathop{\mathfrak{sp}}(n)_{\mathbb C}
 \cong \mathop{\mathfrak{sp}}(n;\mathbb C).
\end{align}
On the other hand, a complex simple Lie algebra $\mathfrak a$ can be, on the other hand, viewed as a real Lie algebra, whose (real) dimension is twice as the original (complex) dimension.
We describe such real Lie algebra by $\mathfrak a_{\mathbb R}$.

@@

\subsection{Mathematical Foundations for Spinors}
\subsubsection{Clifford Algebra}
\paragraph{Quadratic form}
Let~\cite[\S6.3]{Jacobson1} $V$ be an $n$-dimensional vector space over a field $K$. A quadratic form $Q$ on $V$ is defined by
\begin{equation}
\begin{split}
   Q\colon V\to K,\qquad
   &\forall k\in K, v\in V, Q(k v)=k^2Q(v),\\
   &B\colon (v,w)\mapsto Q(v+w)-Q(v)-Q(w)\text{~is linear in $v$ and $w$.}
\end{split}
\end{equation}
The bilinear form $B$ satisfies $B(x,x)=2Q(x)$. As long as the characteristic of $F$ is not two, $Q$ is determined by $B$.

Equivalently, $Q$ is characterized by a symmetric ``matrix'' $\hat Q$ such that $Q(v)=v^\TT \hat Q v$.
Focusing on $K=\RR$, $\hat Q$ is a real symmetric matrix and thus diagonalizable. Sylvester's law of inertia states that there exists a basis such that $\hat Q=\diag(1,1,\cdots,1,-1,-1,\cdots,-1,0,\cdots,0)$, where the number of $+1$ and $-1$ are respectively $p$ and $q$.
The pair $(p,q)$ is unique and is called the signature of $Q$, where $\rank \hat Q=p+q~~(\le n)$.
Meanwhile, if $K$ is algebraically closed such as $\CC$, we can take $\hat Q=\diag(1,1,\cdots,1,0,\cdots,0)$ and the algebra is characterized by $\rank\hat Q$.

Quadratic form induces orthogonality: a basis $\{e_i\}$ of $V$ is called
\begin{equation}
  \text{orthogonal under $Q$}\stackrel{\text{def}}\iff B(e_i,e_j)\propto \delta_{ij}\qquad
  \Bigl[\text{orthonormal under $Q$}\stackrel{\text{def}}\iff B(e_i,e_j)=c_i\delta_{ij}\text{~with~}c_i\in\{2,0,-2\}\Bigr]
\end{equation}
and $(V,Q)$ has an orthogonal basis as long as the characteristic of $K$ is not two.
If $K$ is a ``spin field'' such as $\RR$ or $\CC$ (or, precisely, $\forall \alpha\in K,\exists \beta\in K$ s.t. $\alpha=\beta^2$ or $-\beta^2$), $(V,Q)$ has an orthonormal basis.

\paragraph{Geometric algebra}
Let~\cite{Lundholm:2009xd} $V$ be an $n$-dimensional vector space over a field $K$. Let $Q$ be a quadratic form on $V$.
Tensor algebra on $V$ is defined by
\begin{equation}
  T(V) = \bigoplus_{k} V^{\otimes k}
       = K\oplus V\oplus (V\otimes V)\oplus(V\otimes V\otimes V)\oplus\cdots.
\end{equation}
It has an ideal $K_Q$ generated by $\{v\otimes v - Q(v)\mid v\in V\}$:
\begin{equation}
  K_Q = \left\{\sum_{i} a_i\otimes(v_i\otimes v_i-Q(v_i))\otimes b_i\relmiddle| v_i\in V, a_i,b_i \in T(V)\right\},
\end{equation}
and the geometric algebra $\GeoA(V,Q)\deq T(V)/K_Q$.
The product in $\GeoA(V,Q)$ is written by $vw \deq [v\otimes w]=v\otimes w+K_q$. Particularly, $v^2 = Q(v)\in K$ and $vw+wv=B(v,w)\in K$.

The pseudoscalar $I$ is defined by $I\deq e_1e_2\cdots e_n$ with the basis $\{e_i\}$. For $\GeoA(\RR^{s,t,u})$, $I^2=(-1)^{n(n-1)/2+t}\delta_{0u}$ and $I$ commutes (anti-commutes) with all $e_i$ if $n$ is even (odd).

\paragraph{Clifford Algebra}
For a finite set $X$ and a commutative ring $R$ with unit,
the Clifford algebra $\Cl(X,R,r)$ is defined as the free $R$-module generated by the set $2^X$ of all subset of $X$.
Here, $r\colon X\to R$ is an arbitrary function thought of as a signature.
The Clifford algebra $\mathrm{Cl}$ has
addition $\mathrm{Cl}\times\mathrm{Cl}\to\mathrm{Cl}$,
scalar multiplication $R\times\mathrm{Cl}\to\mathrm{Cl}$, and
product
\begin{equation}
  \mathrm{Cl}\times\mathrm{Cl}\to\mathrm{Cl}\ni AB
  \deq \tau(A,B)A\setdiff B
\end{equation}
defined based on the following map $\tau\colon2^X\times2^X\to R$:
\begin{equation}
  \begin{split}
  &\tau(\{x\},\{x\}) = r(x)\quad\forall x\in X,&
  &\tau(A,B)\tau(A\setdiff B,C)=\tau(A,B\setdiff C)\tau(B,C)\quad\forall A,B,C\in 2^X,\\
  &\tau(\{y\},\{x\}) = -\tau(\{x\},\{y\})\quad\forall x,y\in X,x\neq y,&
  &\tau(\emptyset,A)=\tau(A,\emptyset)=1\quad\forall A\in2^X,\\
  &\tau(A,B)\in\{-1,1\}\quad \text{if}~A\cap B=\emptyset,
  \end{split}
\end{equation}
where $A\setdiff B=(A\cup B)\setminus(B\cap A)$. Namely,
\begin{equation}
  \{x\}\{x\} = r(x)\emptyset,\quad
  \{x\}\{y\}=-\{y\}\{x\},\quad
  \emptyset A=A\emptyset=A,\quad
  (AB)C=A(BC),\quad
  A\cap B=\emptyset\Rightarrow AB=\pm1(A\cup B).
\end{equation}
